---
title: "Time Series Analysis and Modeling of the Healthyverse Packages"
author: "Steven P. Sanderson II, MPH - Data Scientist/IT Manager"
date: "Report Date: `r format(Sys.time(), '%B %d, %Y')`"
output: github_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
source("00_scripts/load_libraries.R")
source("00_scripts/get_data_functions.R")
source("00_scripts/helper_functions.R")
source("00_scripts/data_manipulation_functions.R")
source("00_scripts/ts_decomp.R")
```

## Get Data

```{r get_data}
get_cran_data()
csv_to_rds()
downloads_tbl <- downloads_processed_tbl()

glimpse(downloads_tbl)
```

The last day in the dataset is `r max_cran_datetime()`, the file was birthed on:
`r dl_birth_datetime()`, and is `r hours_since_cran_log_update()` hours old.
`r update_log_message()`

Now that we have our data lets take a look at it using the `skimr` package.

```{r skim_data}
skim(downloads_tbl)
```

We can see that the following columns are missing a lot of data and for us are most
likely not useful anyways, so we will drop them `c(r_version, r_arch, r_os)`

```{r data_trimmed}
data_tbl <- downloads_tbl %>%
    select(-r_version, -r_arch, -r_os)
```

