---
title: "Time Series Analysis and Modeling of the Healthyverse Packages"
author: "Steven P. Sanderson II, MPH - Data Scientist/IT Manager"
date: "Report Date: `r format(Sys.time(), '%B %d, %Y')`"
output: github_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
source("00_scripts/load_libraries.R")
source("00_scripts/get_data_functions.R")
source("00_scripts/helper_functions.R")
source("00_scripts/data_manipulation_functions.R")
source("00_scripts/ts_decomp.R")
```

## Get Data

```{r get_data}
get_cran_data()
csv_to_rds()
downloads_tbl <- downloads_processed_tbl()

glimpse(downloads_tbl)
```

The last day in the dataset is `r max_cran_datetime()`, the file was birthed on:
`r dl_birth_datetime()`, and is `r hours_since_cran_log_update()` hours old.
`r update_log_message()`

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
